---
title: "Insurance Data Analysis"
author: "Chirasthi Thennakoon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Insurance Data Analysis

This is a R counterpart of the same Analysis done in python

Link to the python project : <https://github.com/ChirasthiT/Insurance-Data-Analysis-Prediction-Python>

This project will be done in a R-Markdown file (.Rmd), If not previewed please open in a R environment or in rstudio. HTML output will be knitted.

## Reading and Exploring the dataset

```{r}
# Reading the data set
dataset <- read.csv('insurance.csv')
head(dataset)
```

```{r}
col_names <- colnames(dataset)
col_names
```

```{r}
col_names[ !col_names == "charges"]
```

```{r}
plot(dataset, col = ifelse(dataset$smoker == "yes", "red", "green"))
```

## Transforming the Data set 

```{r}
# Transforming categorical values into numericl values (dummy variables)
head(dataset)

dataset_transformed <- dataset

categorical_col_names = c('sex', 'smoker', 'region')
for (name in categorical_col_names) {
  dataset_transformed[[name]] = as.numeric(factor(dataset_transformed[[name]]))
}
head(dataset_transformed)
```

```{r}
plotting_func <- function(col_name) {
  plot(x = dataset_transformed[[col_name]], y = dataset_transformed$charges)
}
```

```{r}
for (name in col_names) {
  plotting_func(name)
}
```

```{r}
head(dataset_transformed)

plot(
  dataset_transformed$charges,
  col = if_else(condition = dataset_transformed$smoker == 2, true = 'red', false = 'green'),
  sub = "Smoker = red || Non Smoker = green"
)

boxplot(
  dataset_transformed$charges ~ dataset_transformed$smoker,
  col = if_else(condition = dataset_transformed$smoker == 2, true = 'green', false = 'red'),
  horizontal = T
)
```

```{r}
subset(dataset, charges > 50000)
```

## Fitting the Linear Model

```{r}
train_indices <- sample(1:nrow(dataset_transformed), 0.8*nrow(dataset_transformed)) # 80% for training
train_data <- dataset_transformed[train_indices, ]
test_data <- dataset_transformed[-train_indices, ]
```

```{r}
linear_model <- lm(formula = charges ~ age + bmi + children + smoker, data = train_data)
summary(linear_model)
```

```{r}
predictions <- predict(object = linear_model, newdata = test_data)

cor.test(x = predictions, y = test_data$charges)
```

```{r}
# Increase the training sample size
train_indices_1 <- sample(1:nrow(dataset_transformed), 0.8*nrow(dataset_transformed))
test_data_1 <- dataset_transformed[train_indices, ]
predictions_1 <- predict(object = linear_model, test_data_1)
cor.test(predictions_1, test_data_1$charges)
```

We have a accuracy of \<0.8, but the model may be over-fitted. Quadratic model may be more suitable to this data set
